---
title: "beecount-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{beecount-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{printr}
---
# bee.count
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, message=FALSE}
library(bee.count)
#library(xts)
library(printr)
data <- bee.data
original <- bee.data[!is.na(TOTAL_COUNT)]
```

```{r, echo=FALSE,fig.width=7, fig.height=1}
par(mar = c(0,0,0,0))
plot(ts(original$TOTAL_COUNT[1:140]), main = NULL,
     ylab = NULL,
     xlab = NULL)
```

# Data Set

Our dataset was produced by an AI using digital particle image velocimetry or 
DPIV [Willert and Gharib, 1991] to count the level of bee activity outside 
a hive marked R_4_5. These csv files are part of the output of
Dr. Sarbajit Mukherjee during his dissertation at Utah State University. We have
received permission of Dr. Kulyukin to use these files for time series analysis.
The counts correspond to a 28 second period recorded approximately every 
15 minutes throughout the day. The monitors shut-down at night and went off line
several times throughout the season leaving significant coverage gaps.

## bee.data

This library contains an R data file named bee.data. It is a collection of bee motion counts calculated by DPIV and paired with the weather conditions organized by timestamp.

<details>
  <summary>Long Description of bee.data</summary>
```{r, echo=FALSE}
help(bee.data)
```
</details> 

---

The R_4_5 monitor worked reasonable well during the 2017 season. However, the 
data collected from is was heavily right skewed and missing multiple values.

```{r, fig.align='center', fig.width=7, echo=FALSE}
par(mfrow = c(1, 2))
hist(data[!is.na(TOTAL_COUNT), TOTAL_COUNT], breaks = seq(0, 13500, 500),
     main = 'Distribution of Bee Counts', xlab = 'Bee Count')
plot(ts(data$TOTAL_COUNT[1:200]), main = "Data Accounting for Missing Times",
     ylab = 'Count',
     xlab = 'Time (15 Min. Intervals)')
```

---

# Time series
Time series data sets are most often indexed by time and store a single value for
each index. A distinctive feature of time series is that the mean and variance
may often be misleading as there is a significant auto correlation between a 
point and the preceding or following points. This makes linear models 
unsatisfactory and instead the data scientist must use the more poorly 
understood techniques of time series analysis.

---

# Purpose
The purpose of the collected data set and associated tools is to provide a 
frame-work for the data scientist to better understand the processing involved
in preparing a data set for time series analysis and interpreting the results.

---

# Functions

## base_impute

The base_impute function replaces missing values to a value indicated to by the
programmer. 

<details>
  <summary>Long Description of base_impute</summary>
```{r, echo=FALSE}
help(base_impute)
```
</details>

## sparse_ar

The sparse_ar function computes an autoregressive integrated moving average
without the user selecting all the needed parameters of the stats package
arima function.


<details>
  <summary>Long Description of spare_ar</summary>
```{r, echo=FALSE}
help(sparse_ar)
```
</details>

## sparse_arma

The sparse_arma function computes an autoregressive moving average while
limiting the variables needed from the user.

<details>
  <summary>Long Description of sparse_arma</summary>
```{r, echo=FALSE}
help(sparse_arma)
```
</details>

## sparse_ma

The sparse_ma function computes an moving average for the user.


<details>
  <summary>Long Description of sparse_ma</summary>
```{r, echo=FALSE}
help(sparse_ma)
```
</details>

---

# Tutorial


